---
title: "social capital index"
author: "Wesley_Tao"
date: "2018.3.29"
output: html_document

  
---

Task 1: Please provide a visualization of this data in a map format that describes the changing social capital by county from years 1997 to 2014. 
Task 2: Provide analyses on the change of social capital over time by the statistical method of your choice

```{r message=FALSE}
# import the packages we are going to use
library("xlsx") 
library("choroplethr")
library("choroplethrMaps")
# load the data
table_1997<-read.xlsx("../data/social capital 1997-2014.xlsx", 1)
table_2005<-read.xlsx("../data/social capital 1997-2014.xlsx", 2)
table_2009<-read.xlsx("../data/social capital 1997-2014.xlsx", 3)
table_2014<-read.xlsx("../data/social capital 1997-2014.xlsx", 4)
# change the header so we could align dataframe
names(table_2014)<-c("fips","area_name","sk14")
names(table_2009)
names(table_2005)
names(table_1997)
```

# merge data
```{r}
# save them in the output for further exploration
# converge them in the same table and save them in csv file for further exploration
library(dplyr)
m.1<-table_1997 %>% 
  full_join(table_2005, by = c("fips")) %>%
  select(fips, sk97, sk05)

m.2<-table_2009 %>% 
  full_join(table_2014, by = c("fips")) %>%
  select(fips, sk09, sk14)

merged<-m.1 %>%
  full_join(m.2, by = c("fips")) %>%
  select(fips, sk97, sk05, sk09, sk14)
nrow(merged)

write.csv(merged,file="../output/merged_4_year.csv")# store the whole table
```

# store each year of data in tsv file
```{r}
out_97<-table_1997[,c("fips","sk97")]
names(out_97)<-c("id","rate")
write.table(out_97, file="../output/rate_1997.tsv", quote=FALSE, sep='\t', row.names = FALSE)

out_05<-table_2005[,c("fips","sk05")]
names(out_05)<-c("id","rate")
write.table(out_05, file="../output/rate_2005.tsv", quote=FALSE, sep='\t', row.names = FALSE)

out_09<-table_2009[,c("fips","sk09")]
names(out_09)<-c("id","rate")
write.table(out_09, file="../output/rate_2009.tsv", quote=FALSE, sep='\t', row.names = FALSE)

out_14<-table_2014[,c("fips","sk14")]
names(out_14)<-c("id","rate")
write.table(out_14, file="../output/rate_2014.tsv", quote=FALSE, sep='\t', row.names = FALSE)

```

```{r}
library(reshape2)
head(merged)
merged.melt<-melt(merged,id.vars = "fips",variable_name = "year",value_name="index")
merged.melt$year<-as.factor(merged.melt$year)


```


```{r}
library(ggplot2)
# plot the histogram 
ggplot(data=merged.melt,aes(x=value,fill=year))+
  geom_density(alpha=0.3)
```
We do discover a trend that social capital index distribution become more centered. 

```{r}
head(merged)
ggplot(data=merged,aes(x=sk97))+
  geom_histogram()
```
```{r}
ggplot(data=merged,aes(x=sk09))+
  geom_histogram()
```



```{r message=FALSE,warning=FALSE}

# plotmap function encapsulate everything
plotmap<-function(df,year){
  tempdata<-df[,c(1,3)]
  names(tempdata)<-c("region","value")

  county_choropleth(tempdata,
                  title = paste(year, "Social Capital index by county"),
                  legend="index")
 
  }
```


```{r warning=FALSE}
# map for 1997
# save images
# setwd("../figs") # set image path
# png(filename = paste("1997",".png",sep = ""))
plotmap(table_1997,"1997")
# dev.off()
```
```{r warning=FALSE}
setwd("../figs") # set image path
png(filename = paste("1997",".png",sep = ""))
plotmap(table_2005,"2005") # map for 2006
dev.off()
```


```{r warning=FALSE}
# setwd("../figs") # set image path
# png(filename = paste("1997",".png",sep = ""))
plotmap(table_2009,"2009") # map for 2009
# dev.off()
```

```{r warning=FALSE}
# setwd("../figs") # set image path
# png(filename = paste("1997",".png",sep = ""))
plotmap(table_2014,"2014") # map for 2014
# dev.off()
```













