---
title: "Social_Capital_2_Simple_Modeling"
author: "Wesley_Tao"
date: "2018/5/21"
output: html_document
---

```{r}
library(lme4)
library(ggplot2)
library("mediation")
social.capital<-read.csv("../output/merged.csv")
head(social.capital)
```

# Gini Coefficient
## Fixed Effect Model for Gini_coefficient
Question: Are jim_crow states and non_jim_crow states have different Gini_coefficient?
Models: one-way anova
$$ y_{ij}=\mu+\alpha_j+\epsilon_{ij}$$
$y_{ij}$represent the gini coefficient of county i in state j  

$\mu$ stands for the grand mean

$\alpha_{j}$ stands for the dummy variable introduced by jim crow law state

```{r}
simple.one.way<-aov(Gini_Coefficient~Jim_Crow,data=social.capital)
summary(simple.one.way)
```

Answers: 
Null hypothesis for this test is: $$ H_0 : \alpha =0 \space H_a: \alpha \neq 0$$
According to the summary statistics (Pr(>F)<0.0001) of this simple analysis, we conclude that jim_crow_states and non_jim_crow states have very different level of gini_coefficient.



## Random Effect for Gini_coefficient
In the previous model, all the counties are divided into 2 groups and within each group they are treated identical. Now we would like to capture the state variability in our model.



```{r}
this_oder<-order(social.capital$Jim_Crow,social.capital$State)
social.capital<-social.capital[this_oder,]
ggplot(data=social.capital)+
  geom_boxplot(aes(x=reorder(State,Jim_Crow),y=Gini_Coefficient,color=Jim_Crow))+
  labs(title="box plot for gini coefficient for each state")
```

We are going to introduce group specific random effect for each state in the model: which means the intercept for each states are different.
$$ y_{ij}=\alpha_j+\beta_j+\epsilon_{ij}$$
$\beta_j$ stands for the intercept for group j
$\alpha_j$ stands for the jim_crow dummy variable

```{r}
r.e.1<-lmer(Gini_Coefficient~Jim_Crow+(1|State),data=social.capital)
summary(r.e.1)
```
Interpretation: 
states with Jim_Crow label have a marginal increase 0.013 increase in Gini Coefficient


## Statistical significance
We need some sort of p values unfortunately p-values for mixed model are not straight forward as they are in linear models.   
I would like to introduce Likelihood Ratio Test as a means to attain P-values  
Null model: states with varying intercept without jim_crow
Alternative model: with jim_crow

```{r}
gini.null<-lmer(Gini_Coefficient~(1|State),data=social.capital)
gini.alter<-lmer(Gini_Coefficient~Jim_Crow+(1|State),data=social.capital)
anova(gini.null,gini.alter)
```

Conclusion: Jim_crow dummy affect Gini Index ($\chi^2(1)=8.62 p=0.0033$) raising about 0.01 $\pm$ 0.004(standard errors)

# Social Capital index multilevel regression
We are going to do similar statistical test for social capital index for year 1997,2005,2009,2014
## year 1997
```{r}
multilevel.regression<-function(year,df){
  title.name<-paste("Box Plot for ",year," for each state")
  p<-ggplot(data=df,environment = environment())+
  geom_boxplot(aes_string(x="reorder(State,Jim_Crow)",y=year,color="Jim_Crow"))+
  labs(title=title.name)
  
  f.null<-as.formula(paste(year,"~(1|State)"))
  f.alter<-as.formula(paste(year,"~ Jim_Crow+(1|State)"))
  
  gini.null<-lmer(f.null,data=df)
  gini.alter<-lmer(f.alter,data=df)
  print(summary(gini.alter))
  cat("likelihood ratio test for two models")
  print(anova(gini.null,gini.alter))
  return(p)
}
multilevel.regression("sk97",social.capital)
```

# year 2005
```{r}
multilevel.regression("sk05",social.capital)
```

## year 2009

```{r}
multilevel.regression("sk09",social.capital)
```

```{r}
multilevel.regression("sk14",social.capital)
```


# Mediation analysis

Mediation analysis for Jim Crow Laws + Gini Coefficient --> Social Capital (2014)
```{r}
out.fit<-lm(sk14~Jim_Crow+Gini_Coefficient,data=social.capital)
med.fit<-lm(Gini_Coefficient~Jim_Crow,data=social.capital)
med.out<-mediation::mediate(med.fit,out.fit,treat = "Jim_Crow",mediator = "Gini_Coefficient")
summary(med.out)

```
This method is using simulation samples to construct the confidence interval for direct effect and mediate effect.

ADE stands for average direct effect 
$$Direct\space Effect=Y_i(1,M_i(t))-Y_i(0,M_i(t)))$$

ACME means the average causal mediation effects 
$$Mediation\space Effect=Y_i(t,M_i(1))-Y_i(t,M_i(0)))$$

$$ Total\space Effect = Mediation \space Effect + Direct\space Effect $$


prop.mediated stands for the proportion of meidation effect. This is a ratio 

$$ prop.mediate = mediate.effect/direct.effect $$ 



```{r}
cat("Prop Mediated is a ratio of two estimates, which are known to have a very high variance especially when the average treatment effect is not precisely estimated.")
cat(" I would focus on the point estimate of this quantity rather than its CI.  The most important thing is the ACME which have a much more stable CI.")
```


```{r}
plot(med.out)
```

```{r}
cat(" mediation effect is very small compared to direct effect. ")
```
